### 2016-05-02 ch03-lrm/nonparametric.tex
Small stuff in CH03

Whereas many areas of system identification focus on parametric identification and/or parameter estimation, non-parametric techniques have recently received renewed interest from the community.
One of the reasons for this is that non-parametric techniques typically require less assumptions and knowledge of the system than building parametric models.
As a consequence, non-parametric methods are often more robust to wrong assumptions than fully parametric methods.
Besides, \glspl{FRF} have been shown to provide a quick insight into the basic dynamic properties of \gls{LTI} systems.
This capability is very useful for system design, controller design, model validation and model selection for parametric models~\citep{Pintelon2012}.

The measurement of \glspl{FRF} of dynamic systems is an important step in many technical applications, often used as a simple visualization of the system dynamics.
However, obtaining a good non-parametric \gls{FRF} from a measured input-output data set can prove difficult due to the presence of noise, leakage and measurement time limitations to observe lightly-damped dynamics well.
\gls{FRF} measurement techniques are discussed, for instance, in
\citep{Schoukens1998,Schoukens2006LPM,Guillaume1996,Broersen1995,Pintelon2010LPM1,Antoni2007FRF,Pintelon2012}, and applied to practical devices and systems~\citep{Lim2010,Robinson1990,Behjat2010}, among others.

@@

To estimate the local parameters $\theta$  
\begin{equation}
\theta \isdef 
  \begin{bmatrix}
  \theta_A\\ 
  \theta_B\\
  \theta_C
  \end{bmatrix}
\qquad \text{ with }
\theta_A \isdef
\begin{bmatrix}
a_1\\ \vdots\\ a_{\order{A}}
\end{bmatrix}
\text{, }
\theta_B \isdef
\begin{bmatrix}
b_0\\ \vdots\\ b_{\order{B}}
\end{bmatrix}
\text{, and }
\theta_C \isdef
\begin{bmatrix}
c_0\\ \vdots\\ c_{\order{C}}
\end{bmatrix}
\end{equation}
in \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T}, we consider a local window
\begin{equation}
  \LocalWindow[k] 
  \isdef
  \Set{
    \omega_{k+r} 
    | 
    r \in \LocalShifts{k}
  }
\end{equation}
with $\LocalShifts{k} \isdef \Set{-\numel{W}, -\numel{W}+1, \ldots, 0, \ldots, \numel{W}}$
such that the local window around $\omega_k$ consists of $\nWind = 2 \numel{W} + 1$ bins. 
If one denotes the input/output spectra in such a local frequency window $\LocalWindow[k]$ as
\begin{equation}
  \LocalVector{U}_k \isdef 
  \mat{
    U(\omega_{k-\numel{W}})\\
    \vdots\\
    U(\omega_{k})\\
    \vdots\\
    U(\omega_{k+\numel{W}})\\
  }
  \text{ and }
  \LocalVector{Y}_k \isdef 
  \mat{
    Y(\omega_{k-\numel{W}})\\
    \vdots\\
    Y(\omega_{k})\\
    \vdots\\
    Y(\omega_{k+\numel{W}})\\
  }
  \text{,}
\end{equation}
and similarly for the other quantities, equation \eqref{eq:output-error-FD} limited to $\LocalWindow[k]$ can be written as
\begin{equation}
  \LocalVector{Y}_k  = {\true{\LocalVector{G}}}_k \hadamard \LocalVector{U}_k + \LocalVector{T}_k + \LocalVector{V}_k
\text{,}
\end{equation} 
where $\hadamard$ denotes the element-wise product (also, Hadamard product).
Substituting $\true{\LocalVector{G}}$ and $\LocalVector{T}$ with the local models $\LocalModel{G}$ and $\LocalModel{T}$ and neglecting the influence of $\LocalVector{V}$ yields
\begin{equation}
  \LocalVector{Y} 
  \approx 
  \hat{\LocalVector{Y}} 
    =
      \LocalModel{G} \hadamard \LocalVector{U} + \LocalModel{T}
      \text{.}
      \label{eq:local-system-equations}
\end{equation}
Note that this encompasses $\nWind$ complex equations in the $\nth = \order{A} + \order{B} + \order{C} + 2$  unknown complex model parameters ($a_i$, $b_i$ and $c_i$ in \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T}).
Consequently, a necessary condition to compute the model parameters is that
\begin{equation}
  \DOF = \nWind - \nth
       = 2 \numel{W} - \order{A} - \order{B} - \order{C} - 1
       \label{eq:nparam:DOF}
\end{equation}
is positive.
To effectively compute the local model parameters, the equation error in \eqref{eq:local-system-equations} is minimized by formulating a quadratic cost function:
\begin{equation}
  \LocalVector{\CostFunc{\LRIC}} 
    \isdef 
      \left( \LocalVector{Y}  -  \LocalModel{G} \hadamard \LocalVector{U} - \LocalModel{T} \right)^{\HT} 
      \left( \LocalVector{Y}  -  \LocalModel{G} \hadamard \LocalVector{U} - \LocalModel{T} \right)
   \label{eq:costfunc:LRIC}
   \text{.}
\end{equation}
Due to the presence of the denominator of $\LocalModel{G} = \frac{\LocalModel{B}}{\LocalModel{A}}$, the equation error is not linear in the parameters $a_i$ and hence \eqref{eq:costfunc:LRIC} requires starting values and time-consuming iterative optimization schemes to obtain a parameter estimate.
We hence call this method the \gls{LRIC} and denote specific configurations of this method as $\lric{\numel{W},\order{B},\order{A},\order{C}}$.

@@

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\dW,\theta)
  \approx
  \frac{\LocalModel{G}(\dW, \theta)}
           {\LocalModel{A}(\dW, \theta)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{align}
\frac{\Delta \LocalModel{G}(\dw{\omega}{}, \theta)}{\LocalModel{G}(\dw{\omega}{}, \theta)}
&\approx
\LocalModel{A}(\dw{\omega}{}, \theta)^{-1}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \evaluate{\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}}{A}
  \PartialDerivative{\ELSCost[n]{\true\theta}}{\theta_A}^{\TT}
  \\
  &=
  \frac{\sigma_V^2}{\nWind}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \evaluate{\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}}{A}
  \begin{bmatrix}
     \Sum_{\dW} \dW^1 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \deemph{\vdots}\\
     \Sum_{\dW} \dW^{\order{A}} 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \end{bmatrix}
\end{align}
where $\evaluate{\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}}{A}$ denotes the upper left $\order{A}\times\order{A}$ subblock of $\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}$ which pertains to the parameters of $\LocalModel{A}$.
From the leading factor, it can be seen that the bias on the \gls{FRF} is $\bigO{\sigma_V^2}$.
Consequently, one can expect that the bias is inherently linked to the \gls{SNR} of the output signal in the frequency domain.

>>>>

Whereas many areas of system identification focus on parametric identification and/or parameter estimation, non-parametric techniques have received renewed interest from the community over the last decade.
One of the reasons for this is that non-parametric techniques typically require less assumptions and knowledge of the system than building parametric models.
As a consequence, non-parametric methods are often more robust than fully parametric methods.
Besides, \glspl{FRF} have been shown to provide a quick insight into the basic dynamic properties of \gls{LTI} systems.
This capability is very useful for system design, manual controller tuning, model validation and model selection for parametric models~\citep{Pintelon2012}.

The measurement of \glspl{FRF} of dynamic systems is an important step in many technical applications, often used as a simple visualization of the system dynamics.
However, obtaining a good non-parametric \gls{FRF} from a measured input-output data set can be challenging due to the presence of noise, leakage and measurement time limitations to observe lightly-damped dynamics well.
\gls{FRF} measurement techniques are discussed, for instance, in
\citep{Schoukens1998,Schoukens2006LPM,Guillaume1996,Broersen1995,Pintelon2010LPM1,Antoni2007FRF,Pintelon2012}, and applied to practical devices and systems~\citep{Lim2010,Robinson1990,Behjat2010}, among others.

@@

To estimate the local parameters $\theta$  
\begin{equation}
\theta \isdef 
  \begin{bmatrix}
  \theta_A\\ 
  \theta_B\\
  \theta_C
  \end{bmatrix}
\qquad \text{ with }
\theta_A \isdef
\begin{bmatrix}
a_1\\ \vdots\\ a_{\order{A}}
\end{bmatrix}
\text{, }
\theta_B \isdef
\begin{bmatrix}
b_0\\ \vdots\\ b_{\order{B}}
\end{bmatrix}
\text{, and }
\theta_C \isdef
\begin{bmatrix}
c_0\\ \vdots\\ c_{\order{C}}
\end{bmatrix}
\end{equation}
in \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T}, we consider a local window
\begin{equation}
  \LocalWindow[k] 
  \isdef
  \Set{
    \omega_{k+r} 
    | 
    r \in \LocalShifts{k}
  }
\end{equation}
with $\LocalShifts{k} \isdef \Set{-\numel{W}, -\numel{W}+1, \ldots, 0, \ldots, \numel{W}}$
such that the local window around $\omega_k$ consists of $\nWind = 2 \numel{W} + 1$ bins. 
If one denotes the input/output spectra in such a local frequency window $\LocalWindow[k]$ as
\begin{equation}
  \LocalVector{U}_k \isdef 
  \mat{
    U(\omega_{k-\numel{W}})\\
    \vdots\\
    U(\omega_{k})\\
    \vdots\\
    U(\omega_{k+\numel{W}})\\
  }
  \text{ and }
  \LocalVector{Y}_k \isdef 
  \mat{
    Y(\omega_{k-\numel{W}})\\
    \vdots\\
    Y(\omega_{k})\\
    \vdots\\
    Y(\omega_{k+\numel{W}})\\
  }
  \text{,}
\end{equation}
and similarly for the other quantities, equation \eqref{eq:output-error-FD} limited to $\LocalWindow[k]$ can be written as
\begin{equation}
  \LocalVector{Y}_k  = {\LocalVector{G}}_k \hadamard \LocalVector{U}_k + \LocalVector{T}_k + \LocalVector{V}_k
\text{,}
\end{equation} 
where $\hadamard$ denotes the element-wise product (also, Hadamard product).
Substituting $\true{\LocalVector{G}}$ and $\LocalVector{T}$ with the local models $\LocalModel{G}$ and $\LocalModel{T}$ and neglecting the influence of $\LocalVector{V}$ yields
\begin{equation}
  \LocalVector{Y} 
  \approx 
  \hat{\LocalVector{Y}} 
    =
      \LocalModel{G} \hadamard \LocalVector{U} + \LocalModel{T}
      \text{.}
      \label{eq:local-system-equations}
\end{equation}
Note that this encompasses $\nWind$ complex equations in the $\nth = \order{A} + \order{B} + \order{C} + 2$  unknown complex model parameters ($a_i$, $b_i$ and $c_i$ in \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T}).
Consequently, a necessary condition to compute the model parameters is that
\begin{equation}
  \DOF = \nWind - \nth
       = 2 \numel{W} - \order{A} - \order{B} - \order{C} - 1
       \label{eq:nparam:DOF}
\end{equation}
is positive.
To effectively compute the local model parameters, the equation error in \eqref{eq:local-system-equations} is minimized by formulating a quadratic cost function:
\begin{equation}
  \LocalVector{\CostFunc{\LRIC}} 
    \isdef 
      \left( \LocalVector{Y}  -  \LocalModel{G} \hadamard \LocalVector{U} - \LocalModel{T} \right)^{\HT} 
      \left( \LocalVector{Y}  -  \LocalModel{G} \hadamard \LocalVector{U} - \LocalModel{T} \right)
   \label{eq:costfunc:LRIC}
   \text{.}
\end{equation}
Due to the presence of the denominator of $\LocalModel{G} = \frac{\LocalModel{B}}{\LocalModel{A}}$, the equation error is not linear in the parameters $a_i$ and hence \eqref{eq:costfunc:LRIC} requires starting values and time-consuming iterative optimization schemes to obtain a parameter estimate.
We hence call this method the \gls{LRIC} and denote specific configurations of this method as $\lric{\numel{W},\order{B},\order{A},\order{C}}$.

@@

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\dW,\theta)
  \approx
  \frac{\LocalModel{G}(\dW, \theta)}
           {\LocalModel{A}(\dW, \theta)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{align}
\frac{\Delta \LocalModel{G}(\dw{\omega}{}, \theta)}{\LocalModel{G}(\dw{\omega}{}, \theta)}
&\approx
\LocalModel{A}(\dw{\omega}{}, \theta)^{-1}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}_{A}
  \PartialDerivative{\ELSCost[n]{\true\theta}}{\theta_A}^{\TT}
  \\
  &=
  \frac{\sigma_V^2}{\nWind}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}_{A}
  \begin{bmatrix}
     \Sum_{\dW} \dW^1 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \deemph{\vdots}\\
     \Sum_{\dW} \dW^{\order{A}} 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \end{bmatrix}
\end{align}
where $\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}_{A}$ denotes the upper left $\order{A}\times\order{A}$ subblock of $\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}$ which pertains to the parameters of $\LocalModel{A}$.
From the leading factor, it can be seen that the bias on the \gls{FRF} is $\bigO{\sigma_V^2}$.
Consequently, one can expect that the bias is inherently linked to the \gls{SNR} of the output signal in the frequency domain.

### 2016-05-02 ch03-lrm/truncation.tex
Small stuff in CH03

The estimated impulse response function  $\hat g_{{\LPM}\setminus \mathrm{DC}}(t)$ is obtained explicitly via the \gls{IDFT} of the \gls{LPM}-estimate $\hat G_{\LPM}(\omega_k)$ of the \gls{FRF} in accordance with equation \eqref{eq:defDFT}, \emph{viz}:
\begin{align}\label{eq:impRespiDFT}
\hat g_{{\LPM}\setminus \mathrm{DC
}}(t) &= \frac{1}{\sqrt{N}}\sum_{k=1}^{N-1}\hat G_{\LPM}(\omega_k)e^{\frac{j2\pi kt}{N}}
\end{align}
where the estimated \gls{FRF} in the frequency band between  the Nyquist and the sample frequencies is obtained as follows:
\begin{align}
\hat G_{\LPM}(\omega_{N-k}) = \overline{\hat G_{\LPM}(\omega_k)},\ \text{for}\ k=1,\dots,N/2
\end{align}
(with $\conj{\hat G_{\LPM}}$ the complex conjugate of $\hat G_{\LPM}$) to ensure $\hat g_{\LPM}(t)$ to be real.

@@

\end{enumerate}
This procedure is illustrated in \figref{fig:nparam:trunc:nonparametric:truncation:Ftest} for the system described by the following equations, which are relevant to the \gls{LPM} outlined in \secref{sec:nparam:trunc:LPMFRFest}:
\begin{subequations}
\label{eq:trunc:systemODEs}
\begin{align}
y_0(t)  &= 1.5371y_0(t-1)    -0.9025y_0(t-2) + u(t)\\
y(t)       &= y_0(t) + e(t),
\end{align}
\end{subequations}
where $e(t)$ is a white noise sequence, such that the SNR of the output signal is $78.2\unit{dB}$.
The figure shows segments of the estimated impulse response. The last two segments are dominated by the noise, while all samples at $t<80$ have large system contributions. The algorithm outlined above selects $\truncTime = 80$, beyond which all samples are set to zero, resulting in a smoothed \gls{FRF} estimate.

>>>>

The estimated impulse response function  $\hat g_{{\LPM}\setminus \mathrm{DC}}(t)$ is obtained explicitly via the \gls{IDFT} of the \gls{LPM}-estimate $\hat G_{\LPM}(\omega_k)$ of the \gls{FRF} in accordance with equation \eqref{eq:defDFT}, \emph{viz}:
\begin{align}\label{eq:impRespiDFT}
\hat g_{{\LPM}\setminus \mathrm{DC}}(t) 
&= 
\frac{1}{\sqrt{N}}
\sum_{k=1}^{N-1}
\hat G_{\LPM}(\omega_k)e^{\frac{j2\pi kt}{N}}
\end{align}
where the estimated \gls{FRF} in the frequency band between  the Nyquist and the sample frequencies is obtained as follows:
\begin{align}
\hat G_{\LPM}(\omega_{N-k}) = \overline{\hat G_{\LPM}(\omega_k)},\ \text{for}\ k=1,\dots,N/2
\end{align}
(with $\conj{\hat G_{\LPM}}$ the complex conjugate of $\hat G_{\LPM}$) to ensure $\hat g_{\LPM}(t)$ to be real.

@@

\end{enumerate}
This procedure is illustrated in \figref{fig:nparam:trunc:nonparametric:truncation:Ftest} for the system described by the following equations:
\begin{subequations}
\label{eq:trunc:systemODEs}
\begin{align}
y_0(t)  &= 1.5371y_0(t-1)    -0.9025y_0(t-2) + u(t)\\
y(t)       &= y_0(t) + e(t),
\end{align}
\end{subequations}
where $e(t)$ is a white noise sequence, such that the \gls{SNR} of the output signal is $78.2\unit{dB}$.
The figure shows segments of the estimated impulse response. The last two segments are dominated by the noise, while all samples at $t<80$ have large system contributions. The algorithm outlined above selects $\truncTime = 80$, beyond which all samples are set to zero, resulting in a smoothed \gls{FRF} estimate.
